---
title: "Statistical Inference - Course Project Part 1"
author: "José Ignacio Gavara"
date: "23/1/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Overview
The purpose of this data analysis is to investigate the exponential distribution
and compare it to the Central Limit Theorem. 

To achieve our goal, we will follow the following steps:
1. We will show the sample mean and compare it to the theoretical mean of the distribution.
2. We will show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. We will show that the distribution is approximately normal.

For this analysis, the lambda will 
be set to 0.2 for all of the simulations. This investigation will compare the 
distribution of averages of 40 exponentials over 1000 simulations.


## Simulations
First, we will set the simulation variables lambda, exponentials, and seed.
```{r}
ECHO=TRUE
set.seed(25)
lambda = 0.2
exponentials = 40
```

Run Simulations with variables
```{r}
simMeans = NULL
for (i in 1 : 1000) simMeans = c(simMeans, mean(rexp(exponentials, lambda)))
```


## Sample Mean vs Theoretical Mean
#### Sample Mean
Calculating the mean from the simulations with give the sample mean.
```{r}
mean(simMeans)
```
#### Means Histogram
```{r}
hist(simMeans, breaks=40, xlim = c(2,9), main="Exponential Function Simulation Means Distribution", col = "blue")
```

#### Theoretical Mean
The theoretical mean of an exponential distribution is:
lambda^-1.
```{r}
lambda^-1
theoretical.mean<-lambda^-1
```

#### Comparison (sample mean - theoretical mean).
We can see that the simulations sample mean and the 
exponential distribution theoretical mean are very similar.The difference between them is:
```{r}
abs(mean(simMeans)-lambda^-1)
```
In percentage (%): 
```{r}
abs((((mean(simMeans)-(lambda^-1))/(mean(simMeans)))*100))
```

## Sample Variance versus Theoretical Variance
#### Sample Variance
Calculating the variance from the simulation means with give the sample
variance.
```{r}
var(simMeans)
```

#### Theoretical Variance
The theoretical variance of an exponential distribution is: 
(lambda * sqrt(n))^-2.
```{r}
(lambda * sqrt(exponentials))^-2
```

#### Comparison
We can see that the simulations sample variance and 
the exponential distribution theoretical variance are also very similar.The difference between variances is only of:
```{r}
abs(var(simMeans)-(lambda * sqrt(exponentials))^-2)
```
In percentage (%): 
```{r}
abs((((var(simMeans)-(lambda * sqrt(exponentials))^-2))/(var(simMeans)))*100)
```

## Distribution 
This is a density histogram of the 1000 simulations.The blue line represents the best fit curve of the sample means, while the red line represents the normal distribution curve. As we can see, the distribution of means of our sampled exponential distributions is very close to a normal distribution, due to the Central Limit Theorem.If we increased our number of samples the distribution would be even closer to the standard normal distribution.
```{r}
library(ggplot2)
dist<-ggplot(data.frame(y=simMeans), aes(x=y)) + 
  geom_histogram(aes(y=..density..), binwidth=0.2, fill="#0072B2", 
                 color="black") +
  stat_function(fun = dnorm, color = "red", linetype = "dashed", size = 1, 
                  args = list(mean = lambda^-1, sd=(lambda*sqrt(exponentials))^-1), 
                size=2) +
  geom_vline(xintercept=mean(simMeans), colour="blue", linetype="dashed") +   
geom_vline(xintercept=theoretical.mean, colour="red", linetype="dashed") +
  geom_density(alpha=.2, color = "blue") +
  
  labs(title="Simulations vs normal distribution", x="Simulation Mean")
dist
```
